---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

In our analysis we will use ggplot2 for plotting and dplyr for data manipulation.
```{r warning=FALSE, message=FALSE, error=FALSE}
require(ggplot2)
require(dplyr)
```

## Loading and preprocessing the data
1. We load the data using read.csv()
    ```{r echo=TRUE}
    setwd("~/git/RepData_PeerAssessment1")
    unzip('activity.zip')
    data <- read.csv('activity.csv', na.string='NA', 
                    colClasses = c('integer', 'character', 'integer'))
    ```
2. We convert the date column (it should be POSIXct as dplyr doesn't accept POSIXlt, POSIXt)
    ```{r echo=TRUE}
    data$date <- as.POSIXct(data$date)
    ```


## What is mean total number of steps taken per day?
0. Ignore NAs
    ```{r echo=TRUE}
    complete <- data[complete.cases(data),]
    ```
1. We summarise the data using `dplyr` in order to 
calculate the total number of steps taken per day.
    ```{r echo=TRUE, warning=FALSE}
    summarized <- group_by(complete, date) %>% summarise(sum = sum(steps, na.rm=T))
    ```
2. Using our summarized data frame, we can plot 
the histogram of the total number of steps taken each day
    ```{r echo=TRUE}
    qplot(sum, data=summarized, binwidth=1000)
    ```
3. We calculate the mean and median of the total number of steps taken per day
    ```{r echo=TRUE}
    mean(summarized$sum, na.rm=T)
    median(summarized$sum, na.rm=T)
    ```


## What is the average daily activity pattern?
1. We make a time series plot of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)
    ```{r echo=TRUE}
    averaged <- group_by(data, interval) %>% summarise(avg = mean(steps, na.rm=T))
    qplot(interval, avg, data=averaged, geom='line')
    ```
2. Which 5-minute interval, on average across all the days in the dataset, 
contains the maximum number of steps?
    ```{r echo=TRUE}
    averaged$interval[which.max(averaged$avg)]
    ```


## Imputing missing values
1. Calculate and report the total number of missing values 
in the dataset (i.e. the total number of rows with NAs)
    ```{r echo=TRUE}
    sum(1-complete.cases(data))
    ```
2. We fill NAs by filling average value for that 5min interval
    ```{r echo=TRUE}
    nas <- is.na(data$steps)
    filled <- data
    filled$steps[nas] <- averaged$avg[match(data$interval[nas], averaged$interval)]
    ```
4. Now we can plot another histogram
    ```{r echo=TRUE}
    summarized2 <- group_by(filled, date) %>% summarise(sum = sum(steps, na.rm=T))
    qplot(sum, data=summarized2, binwidth=1000)
    mean(summarized2$sum, na.rm=T)
    median(summarized2$sum, na.rm=T)
    ```
In our first dataset (data) there were 8 days full of NAs, those were replaced with mean day
in our second dataset. As a result we can see increase in 10k-11k category (corresponding to mean).
Mean wasn't changed, while median changed to mean and variance was decreased.


## Are there differences in activity patterns between weekdays and weekends?
1. We create factor variable with 2 levels weekday/weekend.
    ```{r echo=TRUE}
    filled$weekend <- 
        factor(weekdays(filled$date) %in% c("Sunday","Saturday"), labels=c('weekday','weekend'))
    ```
2. We make a panel plot containing a time series plot (i.e. type = "l") 
of the 5-minute interval (x-axis) and the average number of steps taken, 
averaged across all weekday days or weekend days (y-axis).
    ```{r echo=TRUE}
    averaged2 <- group_by(filled, interval, weekend) %>% summarise(avg = mean(steps, na.rm=T))
    qplot(interval, avg, data=averaged2, facets=weekend~., geom='line')
    ```



